European or American? A re-examination of the transatlantic influence over the IMF

### 2025-09-24 manuscript/manuscript.qmd
conclusion: update para on hegemony

These findings have several important implications for our understanding of international financial institutions and global economic governance more broadly. Theoretically, they complicate simple narratives of American hegemony in international organisations. While the United States undoubtedly remains the most powerful single actor within the Bretton Woods system, our results suggest that European powers collectively exercise substantial influence over key institutional functions. This finding aligns with observations about the IMF's historical leadership structure—epitomised by the "gentlemen's agreement" reserving the Managing Director position for Europeans—and suggests that informal power arrangements may be more consequential than formal voting shares or veto rights.

>>>>

These findings have several important implications for our understanding of international financial institutions and global economic governance more broadly. Theoretically, these results complicate conventional narratives of uncontested US hegemony in the IMF. While the United States remains the most powerful single actor within the Bretton Woods system, its influence is concentrated in shaping conditionality rather than in determining access to IMF resources. By contrast, European powers collectively wield substantial institutional authority over lending volumes and approvals, reflecting the Fund’s historical leadership structure epitomised by the “gentlemen’s agreement.” Hegemony within the IMF, therefore, is fragmented rather than monolithic, with American primacy tempered by European institutional strength.
